---
title: "Analyzing Female Mice Body Weights"
author: "Ken Wood"
date: "2024-10-06"
output:
  pdf_document: default
  html_document: default
---

### Objective: 

Your mission is to use R to analyze the body weights of female mice on different diets. You'll apply statistical measures to understand trends and variations in the data.

### Data Overview: 

The dataset you'll be working with is about the body weights of female mice on different diets. The .csv file is attached to this assignment. The dataset contains two columns:

1.	Diet: This column indicates the type of diet each mouse was on. For example, it might be 'chow' or other specific diet types.

2.	Bodyweight: This column shows the bodyweight of each mouse in grams.

### Instructions

1.	Central Tendencies of Body Weights

- Goal: Discover key central tendencies in the body weights of the mice.
- Task: Calculate the mean, median, and mode for the body weights.

Action:
```{r}
# Load the dataset into R.

df <- read.csv('female-mice-weights.csv')
df
```

```{r}

# Apply mean(), median(), and a custom function to find the mode for the Bodyweight column.

avg_bodyweight <- mean(df$Bodyweight)
avg_bodyweight

median_bodyweight <- median(df$Bodyweight)
median_bodyweight

# Mode function
mode_function <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

mode_bodyweight <- mode_function(df$Bodyweight)
mode_bodyweight

# Reflection: Write about which measure (mean, median, or mode) you think best represents the bodyweights and why.
```

2.	Understanding Body Weight Variability

- Goal: Explore the variability in the body weights across all diets.
- Task: Calculate the range, variance, and standard deviation for the body weights.

Action:
```{r}
# Use R’s range(), var(), and sd() functions on the Bodyweight column.

range_bodyweight <- range(df$Bodyweight)
range_bodyweight

var_bodyweight <- var(df$Bodyweight)
var_bodyweight

sd_bodyweight <- sd(df$Bodyweight)
sd_bodyweight

# Reflection: Discuss what these measures tell you about the spread of body weights.
```


3.	Overall Data Summary

- Goal: Gain a comprehensive statistical overview of the dataset.
- Task: Use the summary() function in R.

Action:



```{r}
# Run summary() on the entire dataset.

summary(df)

# Reflection: Highlight any standout points from the summary, such as differences in weight across diets.
```
4.	Diet Comparison
	
- Goal: Compare body weights across different diets.
- Task: Group the data by diet and calculate the average bodyweight for each diet group.

Action:
```{r}

# Utilize the dplyr package to group data and calculate mean body weights.

library(dplyr)
df %>% group_by(Diet) %>% summarise(mean_bodyweight = mean(Bodyweight))

# Reflection: Share insights on how diet might influence bodyweight based on your findings.
```



